Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-10 ICPP-19/communications.tex
minor modifications

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in high performance computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with computations in order to exploit the full potential of these systems. In this article, we consider the problem of deciding the order of data transfers between two memory nodes for a set of independent tasks with the objective to minimize the makespan. We prove that with finite memory capacity obtaining the optimal order of data transfers is an NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities, based on the task graph structure, with computations, which runtime uses to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance, tasks are added recursively based on certain sentinel constraints. For such applications, runtime sees a set of independent ready tasks and schedule them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations such that overlap between computations and communications can be maximised. This is the main topic of this article. We prove that order of communications on two memory nodes with the objective of minimizing the makespan is NP-Complete. Our proof is inspired from work by ~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem. The main difference with their and our approach is that they consider a bounded number of tasks can await execution on second machine and buffer is acquired after completing the task on first machine. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and memory is acquired before starting the data transfer on communication resource. We propose different strategies for runtimes such that it can maximize the overlap of computation and communication for kernels running on two different memory nodes. We evaluate our strategies on a cluster of homogeneous nodes. But our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper.

@@

	The outline of the article is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and describe their favorable scenarios. We mainly consider three categories of heuristics: static heuristics, dynamic heuristics and static heuristics with dynamic corrections. We evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:exp}. Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	The problem considered in this article also can be viewed as a flow shop problem. Communication and computation timings of a task can be considered as processing timings on different machines. Johnson has provided scheduling strategies for 2 and 3-machine flow shop problems with infinite memory capacity~\cite{johnson}. 2-machine flow shop problem with finite buffer has been proven NP-Complete by Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}. They consider a bounded number of tasks can await execution on second machine.

@@

		Any schedule $S$ of $I$ having no idle time must start with $K_0$. We
		first show that no other $K_i$ task can be active with $K_0$,
		otherwise we would get idle time on computation channel. Indeed, the
		communication of such a task $K_i$ would end at time at least $b'>3 +
		6x$, but at most two $A_i$ tasks can be computed, and they end at time
		at most $3+2max\{a_i':1\le i\le 3m\} = 3 + 6x$

@@

	We classify our heuristics into mainly three categories. In the first category, the order of all computations and communications is computed in advance and the same order is followed on both resources. While in the second category, the best suited next task is chosen based on different criteria. In the last category, which is based on combination of strategies of the other two categories, we precompute the order and try to follow it as much as possible, but when the precomputed schedule induces idle time then we look for the best suited next task. In all of our strategies (except linear programming based strategy) order of communications and computations, are same.

@@

		\item $decreasing \_ \allowbreak order \_ \allowbreak of \_ \allowbreak computation \_ \allowbreak strategy$ ($DOCPS$): Order is obtained by sorting all tasks in non-increasing order of computation time. 
		\item $increasing \_ \allowbreak order \_ \allowbreak of \_ \allowbreak communication \_ \allowbreak plus \_ \allowbreak computation \_ \allowbreak strategy$ ($IOCCS$): Order is obtained by sorting all tasks in non-decreasing order of sum of communication and computation times.
		\item $decreasing \_ \allowbreak order \_ \allowbreak of \_ \allowbreak communication \_ \allowbreak plus \_ \allowbreak computation \_ \allowbreak strategy$ ($DOCCS$): Order is obtained by sorting all tasks in non-increasing order of sum of communication and computation times.

@@

	In this class of strategy, when communication channel is idle, a task is chosen which satisfies the selection criteria maximally among other tasks. For example, if the selection criteria is to choose a highly compute intensive task, then we calculate the ratio of compute time and communication time for all tasks, and the task with the maximum ratio is selected (ties are broken randomly). As we assume that order on both channels is same, hence order on compute channel will follow the order of communication resource.

	We select a set of tasks which induce minimum idle time on compute channel and satisfy available memory requirement. Then a task is chosen for processing from this set based on the following heuristics.
	\begin{enumerate}[label=\roman*)]
		%%	\item $largest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: We select a set of tasks which induce minimum idle time on compute channel. A task with the largest communication time is chosen from this set. Memory requirement of this task should not be more than presently available memory. 
		%%	\item $smallest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: A task with the smallest communication time is chosen. Memory requirement of this task should not be more than presently available memory. 
		%%	\item $maximum_accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: We select a set of tasks which induce minimum idle time on compute channel. A task with the maximum ratio of compute time to communication time is chosen from this set. Memory requirement of this task should not be more than presently available memory.

		\item $largest \_ \allowbreak communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($LCMR$): A task with the largest communication time is chosen. 
		\item $smallest \_ \allowbreak communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($SCMR$): A task with the smallest communication time is chosen.
		\item $maximum \_ \allowbreak accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($MAMR$): A task with the maximum ratio of compute time to communication time is chosen.
	\end{enumerate}
	\begin{table}[htb]
		\begin{center}

@@

	In this class of strategy, we precompute the order of tasks based on different criteria and then try to follow the same order as much as possible. But when communication resource is idle due to memory requirement of the next task in the obtained order, then we select a task similar to previous strategy. After a task is selected then we update the remaining order without this task. This class of strategy takes advantage of static information in the form of precomputed order and dynamic correction to minimize the idle time due to memory constraint.
	%%%%alg_name="optimal_order_infinite_memory_largest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="optimal_order_infinite_memory_smallest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="optimal_order_infinite_memory_maximum_accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";

	Initial order is obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. When communication resource is idle due to memory requirement of the next task in the order, then we select a set of tasks which induce minimum idle time on compute channel and satisfy memory requirement. We define the following heuristics based on how we select a task from this set. After that order of remaining tasks is updated without this task.

	\begin{enumerate}[label=\roman*)]
		\item $optimal \_ \allowbreak order \_ \allowbreak infinite \_ \allowbreak memory \_ \allowbreak largest \_ \allowbreak communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($OOLCMR$): A task with the largest communication time is chosen from the set.
		\item $optimal \_ \allowbreak order \_ \allowbreak infinite \_ \allowbreak memory \_ \allowbreak smallest \_ \allowbreak communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($OOSCMR$): A task with the smallest communication time is chosen from the set.
		\item $optimal \_ \allowbreak order \_ \allowbreak infinite \_ \allowbreak memory \_ \allowbreak maximum \_ \allowbreak accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($OOMAMR$): A task, whose ratio of compute time to communication time is maximum, is chosen from the set .
	\end{enumerate}

@@

		\end{tikzpicture}
		\caption{ \label{fig:staticOrderDynamicCorrectionsExample} Different static order dynamic corrections heuristic schedules for a task set of Table~\ref{tab:staticOrderDynamicCorrectionsExample} . The $OMIM$ order is $BCDAE$..}
	\end{figure}

@@

	\subsection{Solving Linear Program Iteratively}
	\label{subsec:linearprogrammingformulation}
	We use a mixed integer linear program to obtain the order of data transfers and computations.  $COMP_i$ and $COMM_i$ represent computation and communication timings of task $i$. Memory capacity of the target system is $C$. In the linear program formulation, $s_i$ and $e_i$ (resp. $s'_i$ and $e'_i$) represent start and end time of communication (resp. computation) for task $i$. The formulation contains $i)$ n-1 boolean variables, $a_{ij}$, for each task $i$ to denote the order of $i$ and $j$ on communication resource $ii)$ n-1 boolean variables, $b_{ij}$, for each task $T_i$ to denote the order of $i$ and $j$ on computation resource, and $iii)$ n-1 boolean variables, $c_{ij}$, for each task $i$ to denote the order of $s_i$ and $e'_j$.

@@

	\vspace*{-0.5cm}
	\begin{align*}
		& \text{minimize } l \text{ such that }\\
		\forall i, \quad & \text{computation of task } i \text{ completes:}\\
		& e'_i \leq l\\
		\forall i, \quad & \text{computation stars after communication ends:}\\
		& e_i \leq s'_i\\
		\forall i, \forall j\ne i, \quad & \text{exclusive communication of task } i \text{ on communication channel:}\\
		& e_j \leq s_i +(1-a_{ij})L\\
		& e_i\leq s_j +a_{ij}L\\
		\forall i, \forall j\ne i, \quad & \text{exclusive computation of task } i \text{ on computation channel:}\\
		& e'_j \leq s'_i +(1-b_{ij})L\\
		& e'_i\leq s'_j +b_{ij}L\\
		\forall i, \forall j\ne i, \quad & \text{task } i \text{ respects memory capacity on communication channel:}\\
		& e'_j \leq s_i +(1-c_{ij})L\\
		& s_i< e'_j +c_{ij}L\\
		& \sum_r (a_{ir} - c_{ir})CM(r) + CM(i) \le C\\
		%%\forall i, \forall j\ne i, \quad & \text{task }T_i \text{ respects memory capacity on computation channel:}\\
		%%& s_j < e'_i +(1-d_{ij})L\\
		%%& e'_i\leq s_j +d_{ij}L\\
		%%& \sum_r (d_{ir} - b_{ir})CM(r) + TM(i) \le C\\
	\end{align*}

	We use GLPK solver v4.65 to solve the above formulation. The solver was unable to handle number of variables required to solve MILP at the scale of our interest. Hence, we solve the linear program iteratively for a small subset of size $k=3,4,5$ (solver was unable to handle number of variables required for $k\ge6$ ) and represent the makespan calculated by this heuristic as $lp.k$. At the boundary of two iterations we fixed the event (communication or computation) of an unfinished task who has started before the boundary point and consider other events flexible. We compute various $lp.k$ values for different memory capacities and observe that most of the other heuristics perform better than this heuristic. Hence, we do not include this heuristic for the comparison in Section~{\ref{sec:exp}. Figure~\ref{fig:iterativeLpSolution} shows the performance of different heuristics with MILP based heuristics for various memory capacities of a single trace file.
		%%	\subsection{MILP Based Strategy}
		%%We use GLPK solver v4.65 to solve the MILP formulation described in Section~\ref{subsec:linearprogrammingformulation}. The solver was unable to handle number of variables required to solve MILP at the scale of our interest. Hence, we solve the linear program iteratively for a small subset of size $k=3,4,5$ (solver was unable to handle number of variables required for $k\ge6$ ) and represent the makespan calculated by this heuristic as $lp.k$. At the boundary of two iterations we fixed the event (communication or computation) of an unfinished task who has started before the boundary point and consider other events flexible. We compute various $lp.k$ values for different memory capacities and observe that most of the other heuristics perform better than this heuristic. Figure~\ref{fig:iterativeLpSolution} shows the comparison of different heuristics with MILP based heuristic for various memory capacities of a single trace file.
		\todo[inline]{SKumar: Obtain MILP based makespans for new memory capacities}
		\begin{figure}[htb]
			\includegraphics[scale=0.5]{./results/makespan_with_lp.pdf}
			\caption{Comparision of different heuristics with MILP solution based heuristic.}
			\label{fig:iterativeLpSolution}
		\end{figure}

@@

		%%	\todo[inline]{Machine information and our model}

		We consider a machine called Cascade, available at PNNL, for our experiments. We obtain traces by running two molecular chemistry applications, double precision version of Hartree-Fock (HF) and Coupled Cluster Single Double (CCSD) of NWChem package on 10 nodes of this machine. Each node is composed of 16 Intel Xeon E5-2670 cores. NWChem takes advantages of a Partitioned Global Address Space Programming Model, Global Arrays (GA), to use shared-memory programming APIs on distributed memory computers. GA dedicates one core of each node to handle other cores, hence we can view a node as being composed of 15 computational cores. We use 150 processes for each application and obtain 150 trace files. We run CCSD with Uracil molecules input and HF with SiOSi molecules (for Uracil molecules, HF has very less workload, each processor executes only around 20 tasks, that is why we chose SiOSi input for HF execution). Each process executes around 300-800 tasks. Our data transfer model is quite simple and we consider that all data transfers between each process's local memory and GA memory take the same route. Modeling of different routes of data transfers for the same source-destination pair, bandwidth sharing for different source-destination pairs and network congestion is more challenging and part of our future work. Our model is simple yet it provides insight to the application developers (or runtime system) that in what-order all data transfers are issued for the same source-destination pair to maximise communication-computation overlap. Our model is easily adaptable to any source-destination pair when there is one fixed route between source and destination (such as between CPU and GPU, one copy engine to transfer data from CPU (resp. GPU) to GPU (resp. CPU) ).
		\todo[inline]{More about tiling}
		Both applications mainly perform two types of computations, tensor transpose and tensor contraction. HF expects to specify a tile size and we set it to 100, so that each core can be efficiently utilize. CCSD automatically determines tile sizes at different program points based on the input molecules. Hence, HF operates on almost homogeneous tiles while CCSD on heterogeneous tiles.

		%%	\todo[inline]{Scheduled based on Gilmore-Gomory algorithm and its performance }
		We also consider two other static heuristics for evaluation. The first heuristic is based on an algorithm, proposed by Gilmore and Gomory, to obtain the minimal cost sequence for a set of jobs. In this algorithm, each job has a start and end state and a cost is associated to change the state. In our context, this cost can be seen as non-overlap time of computation for two adjacent tasks. Here is the main idea of this algorithm. Initially partial sequence of jobs is represented by a graph such that their overlap is maximum. Subsequently edges are added to this graph which minimized the total non-overlap cost and connects two components. When all nodes of this graph is connected then an edge interchange mechanism is taken into account to determine the sequence of jobs, which ensures that the sequence has minimal cost. We refer ~\cite{Gilmore-Gomory:1964} for the detailed algorithm and ~\cite{gitworkrepo} for our implementation. This algorithm does not take memory capacity into consideration and only provides the sequence of processing. We applied the same sequence with different memory capacity restrictions for our experiments and call this heuristic \textit{Gilmore-Gomory} ($GG$).

@@

		Figure~\ref{fig:ratio_to_optimal_best_ccsd} shows that best variants of dynamic and static order with dynamic correction strategies achieve similar performance at minimum memory capacity $m_c$. But as memory capacity increases, heterogeneity allows static order with dynamic corrections based strategies to take advantage of static knowledge to get maximum overlap and dynamic correction to select another task in case of memory capacity limitation. Static strategies also start performing better at the end, which indicates that this application has potential for significant communication-computation overlap and pure dynamic strategies are unable to take this information into account while making scheduling decisions.

@@

		To get more insights into the considered applications, we also look in to their workloads. Figure~\ref{fig:ApplicationProperties} exhibits that HF is a communication intensive application and maximum 20\% overlap can be expected in the best scenario. While in CCSD , communications and computations are almost evenly distributed and 40-75\% communication-computation overlap is possible.

>>>>

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in high performance computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with computations in order to exploit the full potential of these systems. In this paper, we consider the problem of deciding the order of data transfers between two memory nodes for a set of independent tasks with the objective to minimize the makespan. We prove that with finite memory capacity obtaining the optimal order of data transfers is an NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities, based on the task graph structure, with computations, which runtime uses to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance, tasks are added recursively based on certain sentinel constraints. For such applications, runtime sees a set of independent ready tasks and schedule them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations such that overlap between computations and communications can be maximised. This is the main topic of this paper. We prove that order of communications on two memory nodes with the objective of minimizing the makespan is NP-Complete. Our proof is inspired from work by ~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem. The main difference with their and our approach is that they consider a bounded number of tasks can await execution on second machine and buffer is acquired after completing the task on first machine. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and memory is acquired before starting the data transfer on communication resource. We propose different strategies for runtimes such that it can maximize the overlap of computations and communications. We evaluate our strategies on a cluster of homogeneous nodes. But our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper.

@@

	The outline of the paper is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and describe their favorable scenarios. We mainly consider three categories of heuristics: static heuristics, dynamic heuristics and static heuristics with dynamic corrections. We evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:exp}. Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	The problem considered in this paper also can be viewed as a flow shop problem. Communication and computation timings of a task can be considered as processing timings on different machines. Johnson has provided scheduling strategies for 2 and 3-machine flow shop problems with infinite memory capacity~\cite{johnson}. 2-machine flow shop problem with finite buffer has been proven NP-Complete by Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}. They consider a bounded number of tasks can await execution on second machine.

@@

		Any schedule $S$ of $I$ having no idle time must start with $K_0$. We
		first show that no other $K_i$ task can be active with $K_0$,
		otherwise we would get idle time on computation resource. Indeed, the
		communication of such a task $K_i$ would end at time at least $b'>3 +
		6x$, but at most two $A_i$ tasks can be computed, and they end at time
		at most $3+2max\{a_i':1\le i\le 3m\} = 3 + 6x$

@@

	We classify our heuristics into mainly three categories. In the first category, the order of all computations and communications is computed in advance and the same order is followed on both resources. While in the second category, the best suited next task is chosen based on different criteria. In the last category, which is based on combination of strategies of the other two categories, we precompute the order and try to follow it as much as possible, but when the precomputed schedule induces idle time then we look for the best suited next task. In all of our strategies (except linear programming based strategy) order of communications and computations is same.

@@

		\item $decreasing \_ \allowbreak order \_ \allowbreak of \_ \allowbreak computation \_ \allowbreak strategy$ ($DOCPS$): Order is obtained by sorting all tasks in non-increasing order of computation time. 
		\item $increasing \_ \allowbreak order \_ \allowbreak of \_ \allowbreak communication \_ \allowbreak plus \_ \allowbreak computation \_ \linebreak[4] \allowbreak strategy$ ($IOCCS$): Order is obtained by sorting all tasks in non-decreasing order of sum of communication and computation times.
		\item $decreasing \_ \allowbreak order \_ \allowbreak of \_ \allowbreak communication \_ \allowbreak plus \_ \allowbreak computation \_ \linebreak[4] \allowbreak strategy$ ($DOCCS$): Order is obtained by sorting all tasks in non-increasing order of sum of communication and computation times.

@@

	In this class of strategy, when communication resource is idle, a task is chosen which satisfies the selection criteria maximally among other tasks. For example, if the selection criteria is to choose a highly compute intensive task, then we calculate the ratio of compute time and communication time for all tasks, and the task with the maximum ratio is selected (ties are broken randomly). As we assume that order on both resources is same, hence order on compute resource follows the order of communication resource.

	We select a set of tasks which induce minimum idle time on compute resource and satisfy available memory requirement. Then a task is chosen for processing from this set based on the following heuristics.
	\begin{enumerate}[label=\roman*)]
		%%	\item $largest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: We select a set of tasks which induce minimum idle time on compute channel. A task with the largest communication time is chosen from this set. Memory requirement of this task should not be more than presently available memory. 
		%%	\item $smallest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: A task with the smallest communication time is chosen. Memory requirement of this task should not be more than presently available memory. 
		%%	\item $maximum_accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$: We select a set of tasks which induce minimum idle time on compute channel. A task with the maximum ratio of compute time to communication time is chosen from this set. Memory requirement of this task should not be more than presently available memory.

		\item $largest \_ \allowbreak communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($LCMR$): A task with the largest communication time is chosen. 
		\item $smallest \_ \allowbreak communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($SCMR$): A task with the smallest communication time is chosen.
		\item $maximum \_ \allowbreak accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ \linebreak[4]($MAMR$): A task with the maximum ratio of compute time to communication time is chosen.
	\end{enumerate}
	\begin{table}[htb]
		\begin{center}

@@

	In this class of strategy, we precompute the order of tasks based on different criteria and then try to follow the same order as much as possible. But when communication resource is idle due to memory requirement of the next task in the obtained order, then we select a task similar to previous strategy. After a task is selected then we update the remaining order without this task. This class of strategy takes advantage of static information in the form of precomputed order and dynamic corrections to minimize the idle time due to memory constraint.
	%%%%alg_name="optimal_order_infinite_memory_largest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="optimal_order_infinite_memory_smallest_communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";
	%%%%string alg_name="optimal_order_infinite_memory_maximum_accelerated \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction";

	Initial order is obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. When communication resource is idle due to memory requirement of the next task in the order, then we select a set of tasks which induce minimum idle time on compute resource and satisfy memory requirement. We define the following heuristics based on how we select a task from this set. After that order of remaining tasks is updated without this task.

	\begin{enumerate}[label=\roman*)]
		\item $optimal \_ \allowbreak order \_ \allowbreak infinite \_ \allowbreak memory \_ \allowbreak largest \_ \allowbreak communication \_ \linebreak[4]\allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($OOLCMR$): A task with the largest communication time is chosen from the set.
		\item $optimal \_ \allowbreak order \_ \allowbreak infinite \_ \allowbreak memory \_ \allowbreak smallest \_ \allowbreak communication \_ \allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($OOSCMR$): A task with the smallest communication time is chosen from the set.
		\item $optimal \_ \allowbreak order \_ \allowbreak infinite \_ \allowbreak memory \_ \allowbreak maximum \_ \allowbreak accelerated \_ \linebreak[4]\allowbreak task \_ \allowbreak respects \_ \allowbreak memory \_ \allowbreak restriction$ ($OOMAMR$): A task, whose ratio of compute time to communication time is maximum, is chosen from the set .
	\end{enumerate}

@@

		\end{tikzpicture}
		\caption{ \label{fig:staticOrderDynamicCorrectionsExample} Different static order dynamic corrections heuristic schedules for a task set of Table~\ref{tab:staticOrderDynamicCorrectionsExample} . The $OMIM$ order is $BCDAE$.}
	\end{figure}

@@

	\subsection{Solving Linear Program Iteratively}
	\label{subsec:linearprogrammingformulation}
	We use a mixed integer linear program to obtain the order of communications and computations.  $COMP_i$ and $COMM_i$ represent computation and communication timings of task $i$. Memory capacity of the target system is $C$. In the linear program formulation, $s_i$ and $e_i$ (resp. $s'_i$ and $e'_i$) represent start and end time of communication (resp. computation) for task $i$. The formulation contains $i)$ n-1 boolean variables, $a_{ij}$, for each task $i$ to denote the order of $i$ and $j$ on communication resource $ii)$ n-1 boolean variables, $b_{ij}$, for each task $T_i$ to denote the order of $i$ and $j$ on computation resource, and $iii)$ n-1 boolean variables, $c_{ij}$, for each task $i$ to denote the order of $s_i$ and $e'_j$.

@@

	\vspace*{-0.5cm}
	\begin{align*}
		& \text{minimize } l \text{ such that }\\
		\forall i, \quad & \text{computation of task } i \text{ completes:}\\
		& e'_i \leq l\\
		\forall i, \quad & \text{computation stars after communication ends:}\\
		& e_i \leq s'_i\\
		\forall i, \forall j\ne i, \quad & \text{exclusive communication of task } i \text{ on communication resource:}\\
		& e_j \leq s_i +(1-a_{ij})L\\
		& e_i\leq s_j +a_{ij}L\\
		\forall i, \forall j\ne i, \quad & \text{exclusive computation of task } i \text{ on computation resource:}\\
		& e'_j \leq s'_i +(1-b_{ij})L\\
		& e'_i\leq s'_j +b_{ij}L\\
		\forall i, \forall j\ne i, \quad & \text{task } i \text{ respects memory capacity on communication channel:}\\
		& e'_j \leq s_i +(1-c_{ij})L\\
		& s_i< e'_j +c_{ij}L\\
		& \sum_r (a_{ir} - c_{ir})CM(r) + CM(i) \le C\\
		%%\forall i, \forall j\ne i, \quad & \text{task }T_i \text{ respects memory capacity on computation resource:}\\
		%%& s_j < e'_i +(1-d_{ij})L\\
		%%& e'_i\leq s_j +d_{ij}L\\
		%%& \sum_r (d_{ir} - b_{ir})CM(r) + TM(i) \le C\\
	\end{align*}

	We use GLPK solver v4.65 to solve the above formulation. The solver was unable to handle number of variables required to solve MILP at the scale of our interest. Hence, we solve the linear program iteratively for a small subset of size $k=3,4,5$ (solver was unable to handle number of variables required for $k\ge6$ ) and represent the makespan calculated by this heuristic as $lp.k$. At the boundary of two iterations we fixed the event (communication or computation) of an unfinished task who has started before the boundary point and consider other events flexible. We compute various $lp.k$ values for different memory capacities and observe that most of the other heuristics perform better than this heuristic. Hence, we do not include this heuristic for the comparison in Section~{\ref{sec:exp}. Performance comparison of different heuristics with MILP based heuristics is available in an extended version(cite).

		\todo[inline]{SKumar: Move to an extended version, Obtain MILP based makespans for new memory capacities}
 Figure~\ref{fig:iterativeLpSolution} shows the performance of different heuristics with MILP based heuristics for various memory capacities of a single trace file.
		%%	\subsection{MILP Based Strategy}
		%%We use GLPK solver v4.65 to solve the MILP formulation described in Section~\ref{subsec:linearprogrammingformulation}. The solver was unable to handle number of variables required to solve MILP at the scale of our interest. Hence, we solve the linear program iteratively for a small subset of size $k=3,4,5$ (solver was unable to handle number of variables required for $k\ge6$ ) and represent the makespan calculated by this heuristic as $lp.k$. At the boundary of two iterations we fixed the event (communication or computation) of an unfinished task who has started before the boundary point and consider other events flexible. We compute various $lp.k$ values for different memory capacities and observe that most of the other heuristics perform better than this heuristic. Figure~\ref{fig:iterativeLpSolution} shows the comparison of different heuristics with MILP based heuristic for various memory capacities of a single trace file.
		\begin{figure}[htb]
			\includegraphics[scale=0.5]{./results/makespan_with_lp.pdf}
			\caption{Comparision of different heuristics with MILP solution based heuristic.}
			\label{fig:iterativeLpSolution}
		\end{figure}

@@

		%%	\todo[inline]{Machine information and our model}
		\subsection{Experimental Settings}

		We consider a machine called Cascade, available at PNNL, for our experiments. We obtain traces by running two molecular chemistry applications, double precision version of Hartree-Fock (HF) and Coupled Cluster Single Double (CCSD) of NWChem package on 10 nodes of this machine. Each node is composed of 16 Intel Xeon E5-2670 cores. NWChem takes advantages of a Partitioned Global Address Space Programming Model, Global Arrays (GA), to use shared-memory programming APIs on distributed memory computers. GA dedicates one core of each node to handle other cores, hence we can view a node as being composed of 15 computational cores. We use 150 processes for each application and obtain 150 trace files. We run CCSD with Uracil molecules input and HF with SiOSi molecules (for Uracil molecules, HF has very less workload, each processor executes only around 20 tasks, that is why we chose SiOSi input for HF execution). Each process executes around 300-800 tasks. Our data transfer model is quite simple and we consider that all data transfers between each process's local memory and GA memory take the same route. Modeling of different routes of data transfers for the same source-destination pair, bandwidth sharing for different source-destination pairs and network congestion is more challenging and part of our future work. Our model is simple yet it provides insight to the application developers (or runtime system) that in what-order all data transfers are issued for the same source-destination pair to maximise communication-computation overlap. Our model is easily adaptable to any source-destination pair when there is one fixed route between source and destination (such as between CPU and GPU, one copy engine to transfer data from CPU (resp. GPU) to GPU (resp. CPU) ).
		\todo[inline]{More about tiling}
		Both applications mainly perform two types of computations, tensor transpose and tensor contraction. HF expects to specify a tile size and we set it to 100, so that each core can be efficiently utilize. CCSD automatically determines tile sizes at different program points based on the input molecules. Hence, HF operates on almost less heterogeneous tiles while CCSD on more heterogeneous tiles.

		%%	\todo[inline]{Scheduled based on Gilmore-Gomory algorithm and its performance }
		We also consider two other static heuristics for evaluation. The first heuristic is based on an algorithm, proposed by Gilmore and Gomory, to obtain the minimal cost sequence for a set of jobs~\cite{Gilmore-Gomory:1964}. This is classic in 2-machine no-wait flow shop. In this algorithm, each job has a start and end state and a cost is associated to change the state. In our context, this cost can be seen as non-overlap time of computation for two adjacent tasks. Here is the main idea of this algorithm. Initially partial sequence of jobs is represented by a graph such that their overlap is maximum. Subsequently edges are added to this graph which minimized the total non-overlap cost and connects two components. When all nodes of this graph is connected then an edge interchange mechanism is taken into account to determine the sequence of jobs, which ensures that the sequence has minimal cost. We refer ~\cite{Gilmore-Gomory:1964} for the detailed algorithm and ~\cite{gitworkrepo} for our implementation. This algorithm does not take memory capacity into consideration and only provides the sequence of processing. We applied the same sequence with different memory capacity restrictions for our experiments and call this heuristic \textit{Gilmore-Gomory} ($GG$).

@@

		Figure~\ref{fig:ratio_to_optimal_best_ccsd} shows that best variants of dynamic and static order with dynamic corrections strategies achieve similar performance at minimum memory capacity $m_c$. But as memory capacity increases, heterogeneity allows static order with dynamic corrections based strategies to take advantage of static knowledge to get maximum overlap and dynamic correction to select another task in case of memory capacity limitation. Static strategies also start performing better at the end, which indicates that this application has potential for significant communication-computation overlap and pure dynamic strategies are unable to take this information into account while making scheduling decisions.

@@

		To get more insights into the considered applications, we also look in to their sum of communication timings ($sum_comm$)  and sum of computation timings  ($sum_comp$) with respect to $OMIM$ in Figure~\ref{fig:ApplicationProperties}. It exhibits that HF is a communication intensive application and maximum 20\% overlap can be expected in the best scenario. While in CCSD , communications and computations are almost evenly distributed and 40-75\% communication-computation overlap is possible.
